---
title: "CO2 2030 Gridmap"
output: html_document
date: "2024-10-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r cars}
library(rnaturalearth)
library(sf)
library(ggplot2)
library(openxlsx)
library(dplyr)
```


## Load EDGAR data for 2022
```{r}
emissions_data <- read.table("v8.0_FT2022_GHG_GWP_100_AR5_GHG_2022_TOTALS.txt", header = TRUE, sep = ";")
```

## Create world map and match EDGAR data to the map
```{r pressure, echo=FALSE}
# Load the country boundaries data
world <- ne_countries(scale = "medium", returnclass = "sf")

# Convert the emissions data frame to a spatial object
r_sf <- st_as_sf(emissions_data, coords = c("lon", "lat"), crs = 4326)

# Perform a spatial join to assign countries to each grid cell
grid_with_countries <- st_join(r_sf, world[, c("iso_a3", "name")], join = st_intersects)

grid_with_countries$x <- st_coordinates(grid_with_countries)[, 1]
grid_with_countries$y <- st_coordinates(grid_with_countries)[, 2]
```

## Load our emissions by country data
This is where our ratios for each country are
```{r}
world_totals <- read.xlsx("World Totals New.xlsx")
```

## Merge the spacial data with the emissions by country
```{r}
# Merge the country ratios with the grid data
grid_with_ratios <- merge(grid_with_countries, world_totals, by.x = "name",  by.y = "PARTY", all.x = TRUE)
```

## Apply each countries ratio to the EDGAR data
```{r}
# Apply the ratio to the emissions data, adjust the values
grid_with_ratios$Adjusted_Emissions <- grid_with_ratios$emission * grid_with_ratios$Ratio
```

```{r}
# Summarize emissions for each grid cell
grid_summary <- aggregate(Adjusted_Emissions ~ x + y, data = grid_with_ratios, FUN = sum, na.rm = TRUE)
```

## Rename
```{r}
colnames(grid_summary)[colnames(grid_summary) == "x"] <- "lon"
colnames(grid_summary)[colnames(grid_summary) == "y"] <- "lat"
```

## Save the new data as a CSV
```{r}
write.csv(grid_summary, "Adjusted Emissions (1).csv", row.names = FALSE)
```

## Test for correct emission totals
```{r}
total_emissions_by_country <- grid_with_ratios %>%
  filter(name == "Afghanistan") %>%
  summarise(total_emissions = sum(`Adjusted_Emissions`, na.rm = TRUE))

total_emissions_by_country
```

## Load Adjusted Emissions Data
```{r}
adjusted_data <- read.csv("Adjusted Emissions (1).csv")
```

## Create world map and match data to the map
```{r pressure, echo=FALSE}
# Load the country boundaries data
world <- ne_countries(scale = "medium", returnclass = "sf")

# Convert the emissions data frame to a spatial object
r_sf <- st_as_sf(adjusted_data, coords = c("lon", "lat"), crs = 4326)

# Perform a spatial join to assign countries to each grid cell
grid_with_countries <- st_join(r_sf, world[, c("iso_a3", "name")], join = st_intersects)

grid_with_countries$x <- st_coordinates(grid_with_countries)[, 1]
grid_with_countries$y <- st_coordinates(grid_with_countries)[, 2]
```

## Merge the spacial data with the emissions by country
```{r}
# Merge the country ratios with the grid data
grid_with_ratios <- merge(grid_with_countries, world_data, by.x = "name",  by.y = "PARTY", all.x = TRUE)
```

## Apply each countries CO2 ratio to the data
```{r}
# Apply the ratio to the emissions data, adjust the values
grid_with_ratios$Adjusted_Emissions <- grid_with_ratios$Adjusted_Emissions * grid_with_ratios$ratio.co2
```

```{r}
# Summarize emissions for each grid cell
grid_summary <- aggregate(Adjusted_Emissions ~ x + y, data = grid_with_ratios, FUN = sum, na.rm = TRUE)
```

## Rename
```{r}
colnames(grid_summary)[colnames(grid_summary) == "x"] <- "lon"
colnames(grid_summary)[colnames(grid_summary) == "y"] <- "lat"
```

## Save the new data as a CSV
```{r}
write.csv(grid_summary, "Adjusted Emissions CO2.csv", row.names = FALSE)
```

## Load new data
```{r}
grid_summary <- read.csv("Adjusted Emissions CO2.csv")
```

# Plot Emissions Gridmap
```{r}
ggplot() +
  borders("world", colour = "gray85", fill = "gray80") +
  geom_tile(data = grid_summary, aes(x = lon, y = lat, fill = Adjusted_Emissions)) +
  scale_fill_viridis_c(option = "plasma", name = "Projected 2030 CO2 Emissions (tonnes per 0.1x0.1 degree)", trans = "log10") +
  scale_x_continuous(breaks = seq(floor(min(grid_summary$lon)), ceiling(max(grid_summary$lon)), by = 30)) +
  scale_y_continuous(breaks = seq(floor(min(grid_summary$lat)), ceiling(max(grid_summary$lat)), by = 30)) +
  coord_equal() +
  theme_minimal() +
  theme(
    legend.title = element_text(size = 7), # Customize legend title
  ) +
  labs(title = "Adjusted Global Emissions Map (Logarithmic Scale)", x = "Longitude", y = "Latitude")
```



